---
title: "Getting started"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This document covers the basic use of the package `inflatably`.

### Background

When using costs in a cost analysis (e.g. cost-effectiveness calculation) there are several necessary considerations.

  * Currency
  * Time
  * Context

Often the specific cost with respect to time (and currency) is not available and so we need to inflate the observed cost to the desired target value for a given context.

  * Currency A $\rightarrow$ currency B
  * Time A $\rightarrow$ time B

We fill focus on context-dependent inflation. For further discussion and a focus on multiple currency inflation see [@Turner2019].

For original cost $C_0$, target cost $C_T$ and fraction yearly inflationary factor $i$ for $n$ years, the general inflation formula is as follows.

$$
C_T = (1 + i_{1})(1 + i_{2}) \cdots (1 + i_{n}) \times C_0
$$

Clearly, for the same $i$ each year this simplifies to

$$
C_T = C_0(1 + i)^n
$$

The values of $i$ are context dependent, e.g. different for salaries or resources.
So, to be more precise, we can explicitly denote the inflation factor as $i^s_{y_j}$ where $s$ indicates the source (corresponding to the context) and $y_j$ is the year from which a inflation factor is taken.

$$
C_T = (1 + i^s_{y_1})(1 + i^s_{y_2}) \cdots (1 + i^s_{y_n}) \times C_0
$$


### Example

In this example we inflate the original cost £1 observed in 2014 to 2016 prices using the Hospital and Community Health Services (HCHS) inflation rates.

```{r setup}
library(inflatably)

inflation_adjust_cost_data(from_year = 2014,
                           to_year = 2016,
                           from_cost = 1,
                           inflation_df_nm = "HCHS")
1*1.004*1.01
```

The function returns the inflated cost and also records the inflation source, original year and cost and target year for reproducability.

### References